---
title: '***stringr:*** a package used to manipulate strings'
author: "Devin Teran - Extended By :: Amit Kapoor"
date: "4/16/2020"
output: rmdformats::readthedown

---
### **Getting started**

First we need to load these packages:

* **tidyverse** 
* **stringr**  
* **dplyr** - *used for subsetting data in our analysis*  
* **rmdformats** - *used to for styling html document*   

```{r necessary-libraries,include = FALSE}
packages <- c('tidyverse','stringr','dpylr','rmdformats')

library(tidyverse)
library(stringr)
library(rmdformats)
library(dplyr)
```
  
We're going to load a dataset from [fivethirtyeight.com](https://fivethirtyeight.com/features/a-handful-of-cities-are-driving-2016s-rise-in-murders/) to help us show examples of stringr at work. Our data shows murders in cities in America from 2014 to 2015. 

We'll take the first 10 rows of the data for simplicity's sake.
```{r setup}
url <- 'https://raw.githubusercontent.com/fivethirtyeight/data/master/murder_2016/murder_2015_final.csv'
murder_raw <- read_csv(url)

murder <- head(murder_raw,10)
```

## **Ordering Strings**
#### **str_order(character vector,decreasing = X)**  

**Purpose:**   
Order a character vector alphabetically.  

**Input: **      
character vector - what you want to order  
X - indicate whether to order characters decreasing (FALSE - alphabetically) or increasing (TRUE - order from Z to A)  

**Output:**  
An ordered character vector 

**Example:**  
We'll order the column 'city' from our dataframe 'murder'
```{r sort-descreasing-alphabetically}
str_sort(murder$city,decreasing=FALSE)
```
If you want to reverse the order to Z-A you can set **decreasing = FALSE**
```{r sort-reverse}
str_sort(murder$city,decreasing=TRUE)
```
   
## **Combining Strings**
#### **str_c(String1,String2,...Stringn)**  

**Purpose:**    
The function takes in a strings or vectors of strings and concatentates them together
  
**Input:**     
String or vector of strings separated by comma  
  
**Output:**  
Single string of vector of combined strings  
  
**Example:**  
You can combine as many strings as you want together at once  
```{r combine-many-string}
str_c('a','b','c','d','e','f','g','h')
```  
Let's let's see how we can combine two vectors of strings together from our dataframe: the city and the state  
```{r combine-city-state}
str_c(murder$city,murder$state)
```
You can add a separator between the strings you're combining using the **sep = ''** argument. Let's separate the city and state by a comma.  
  
Add this new data as a column, named **City_State**, in our dataframe murder.
```{r combine-string-separated}
murder$City_State <- str_c(murder$city,murder$state,sep=",")
murder$City_State
```
 
## **Replacing Strings** 
#### **str_replace_all(string, pattern, string)**
**Purpose:**    
This function will replace all instances of a pattern with the given replacement  

**Input:**      
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here

**Output:**  
Single string of vector of combined strings  

**Example:**  
Supposed we wanted to replace all appearances of **.** in the column 'City_State'.  We can easily do this with **str_replace_all()** 
```{r string-replace-all}
murder$City_State <- str_replace_all(murder$City_State,'[\\.]','*')
murder$City_State
```
   
## **Get the Length of a String**
#### **str_length(string)**
**Purpose:**    
Find out the length of a string or a vector of strings   
  
**Input:**       
String or vector of strings     

**Output:**     
Integer    
  
**Example:**  
Let's find how out how long each city name   
```{r string-length}
str_length(murder$city)
```
  
*Let's only view the rows in the dataframe where the city has more than 9 letters in the name.  To do this we'll also use the filter function from the package dplyr.*  
```{r view-more-9-letter-city}
filter(murder,str_length(murder$city) > 9)
```

### **Conclusion**
These examples are just the beginning of what you can do with stringr.  If you need to manipulate, combine or work with strings in general, stringr is a great package to do so.  Here's a great stringr cheatsheet released by RStudio (https://rstudio.com/resources/cheatsheets/).



**Resources:**

* [https://cran.r-project.org/web/packages/rmdformats/vignettes/introduction.html](https://cran.r-project.org/web/packages/rmdformats/vignettes/introduction.html)
* [https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html)



# **Extended By Amit Kapoor**


stringr is built on top of stringi, which provides accurate and fast string manipulation tasks. stringr has most frequently used string manipulation functions while stringi provides a comprehensive set. stringr functions start with str_ prefix and simplifies the manipulation of character strings in R. 

```{r}

# lets see the dataset first before applying stringr functions.
head(murder)
```


## str_which()
### **str_which(string, pattern)**

**Purpose:**    
Identify the location of the character strings containing a certain pattern. It Returns the index of entries that contain the pattern.
  
**Input:**       
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here   

**Output:**     
A vector of integer   
  
**Example:** 


We’ll check the column 'City_State' from our dataframe ‘murder’ to check if it follows the pattern 'alphabets,alphabets' pattern. 

```{r}

# check the patterns alphabets(1 or more times),alphabets(1 or more times)
str_which(murder$City_State, "[a-zA-Z ]+,[a-zA-Z ]+")
```

Here [a-zA-Z ] represents all lower and upper case alphabets and '+' represnts one or more.


## str_match()
### **str_match(string, pattern)**

**Purpose:**    
Extract first part of the string that matches the groups and the patterns defined by the groups.   
  
**Input:**       
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here   

**Output:**     
Character matrix with with one column for the complete match and one column for each group.    
  
**Example:** 

We’ll use the column 'City_State' from our dataframe ‘murder’ to identify the groups.

```{r}
# matches the given pattern.
str_match(murder$City_State, "([a-zA-Z ]+)(,)([a-zA-z ]+)")
```

As described in output, the output shows the first column of complete match and rest 3 for each group of patteren specified.


## str_pad()
### **str_pad(string, width, side = c("left", "right","both"), pad = " ")**

**Purpose:**    
Adds padding characters (default white space) to string to make it a fixed size.   
  
**Input:**       
String or vector of strings
width - Minimum width of padded strings.
side - Side on which padding character is added (left, right or both).
pad - Single padding character (default is a space).   

**Output:**     
String or vector of strings    
  
**Example:** 


Let's use the column 'city' of dataframe 'murder' to demonstrate the padding here. Default padding is on left.

```{r}

# pads '*' on the left in all column values
str_pad(murder$city, 25, pad="*")
```

We can use vector as well for padding.

```{r}

# pads '-' and ' ' on the left in all column values alternatively.
str_pad(murder$city, 15, pad=c("-", " "))
```


## str_dup()

### **str_dup(string, times)**

**Purpose:**    
Repeat a string.   
  
**Input:**       
String or vector of strings
times - Number of times to duplicate each string.

**Output:**     
String or vector of strings  
  
**Example:** 


We will use 'state' here to show the str_dup use. In this example duplicated twice.

```{r}

# Duplicates twice
str_dup(murder$state, 2)
```

Another way of alternate twice and thrice replication.

```{r}

# alternatively repeats twice and thrice
str_dup(murder$state, c(2,3))
```

We can use range as well as described in below example.


```{r}

# replicates as per range provided
str_dup(murder$state, 1:10)
```


## str_detect()
### **str_detect(string, pattern)**

**Purpose:**    
Detects pattern in the string.   
  
**Input:**       
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here   

**Output:**     
boolean or vector of booleans   
  
**Example:** 


We will again use 'City_State' column of given dataframe here to detect the pattern here. It detects the presence or absence of a pattern and returns a logical vector.

```{r}

# returns TRUE based on pattern detected.
str_detect(murder$City_State, "[a-zA-Z ]+,[a-zA-z ]+")
```

If we change the pattern which doesnt match, It returns false. Here I introduced a space after comma.

```{r}


# returns TRUE based on pattern not detected.
str_detect(murder$City_State, "[a-zA-Z ]+, [a-zA-z ]+")
```


## str_locate()
### **str_locate(string, pattern)**

**Purpose:**    
Locates the first position of a pattern and returns a numeric matrix with columns start and end.  
  
**Input:**       
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here  


**Output:**     
A numeric matrix. 
  
**Example:** 


Here we use the column City_State to locate the given pattern. As we can see every value in the given column follows the pattern and the result, locates the the start and end of the pattern.

```{r}

# locates start and end location of given pattern 
str_locate(murder$City_State, "[a-zA-Z ]+,[a-zA-z ]+")
```


## Conclusion
`stringr` package is extrmely useful package for string manipulation which is of great use in data cleaning. It works great with regex patterns as demonstrated above and part of tidyverse package. All functions deal with "NA"'s and zero length vectors in the same way, and the output from one function is easy to feed into the input of another. It is very helpful in data presentation with data of large size.





