---
title: "Data607 - Tidyverse Create Assignment"
author: "Amit Kapoor"
date: "3/24/2020"
output:
  pdf_document: default
  html_document: default
---

### Assignment

Using one or more TidyVerse packages and any dataset from fivethirtyeight.com or Kaggle, the task is to create a programming sample “vignette” that demonstrates use of one or more of the capabilities of selected TidyVerse package.


### Getting started

Lets load tidyverse package first. It includes readr, dplyr, tidyr, ggplot2, stringr, tibble, forcats and purr packages.

* **tidyverse** 


```{r necessary-libraries,include = FALSE}
library(tidyverse)
```
  
We're going to load a dataset from [fivethirtyeight.com](https://fivethirtyeight.com/features/which-state-has-the-worst-drivers/) to help us show tidyverse package at work. This data shows America’s bad drivers in all the states, involved in collisions. 


First step is to read the bad-drivers data from github repository. The data contains below fields:

* State	
* Number of drivers involved in fatal collisions per billion miles
* Percentage Of Drivers Involved In Fatal Collisions Who Were Speeding	
* Percentage Of Drivers Involved In Fatal Collisions Who Were Alcohol-Impaired	
* Percentage Of Drivers Involved In Fatal Collisions Who Were Not Distracted	
* Percentage Of Drivers Involved In Fatal Collisions Who Had Not Been Involved In Any Previous Accidents	
* Car Insurance Premiums ($)	
* Losses incurred by insurance companies for collisions per insured driver ($)



#### Data read using readr package

* **read_csv()** function is from **readr** package, used for reading flat file data with comma separated values.


```{r getdata, echo=TRUE}

# define URL for bad drivers data
theURL <- 'https://raw.githubusercontent.com/fivethirtyeight/data/master/bad-drivers/bad-drivers.csv'

# read data
bad_drivers <- read_csv(theURL)

head(bad_drivers)
```

In the next, we rename columns to replace big column names with shorter names.

* **glimpse()** function is from **tibble** package, used to see every column in a data frame.

```{r rename, echo=TRUE}
# rename columns
colnames(bad_drivers) <- c("STATE", 
                           "DRIVERS_INVOLVED", 
                           "PERC_DRIVERS_SPEED", 
                           "PERC_DRIVERS_ALCHO", 
                           "PERC_DRIVERS_NOT_DIST", 
                           "PERC_DRIVERS_NO_ACC", 
                           "INS_PREM", 
                           "LOSS_INSCOMP")

glimpse(bad_drivers)
```



#### Data wrangling/visualization using dplyr, tidyr and ggplot2 packages

As we must have noticed columns PERC_DRIVERS_SPEED, PERC_DRIVERS_ALCHO, PERC_DRIVERS_NOT_DIST, PERC_DRIVERS_NO_ACC are percentages of DRIVERS_INVOLVED. In the next step we will mutate new columns DRIVERS_SPEED, DRIVERS_ALCHO, DRIVERS_NOT_DIST, DRIVERS_NO_ACC by taking the given percentage of DRIVERS_INVOLVED column.

* **mutate()** function is from **dplyr** package, adds new variables and preserves existing ones.



```{r}
# create new column DRIVERS_SPEED which will be (DRIVERS_INVOLVED*PERC_DRIVERS_SPEED)/100
bad_drivers <- bad_drivers %>% 
  mutate(DRIVERS_SPEED=(DRIVERS_INVOLVED*PERC_DRIVERS_SPEED)/100) %>% 
  mutate(DRIVERS_ALCHO=(DRIVERS_INVOLVED*PERC_DRIVERS_ALCHO)/100) %>% 
  mutate(DRIVERS_NOT_DIST=(DRIVERS_INVOLVED*PERC_DRIVERS_NOT_DIST)/100) %>% 
  mutate(DRIVERS_NO_ACC=(DRIVERS_INVOLVED*PERC_DRIVERS_NO_ACC)/100)

glimpse(bad_drivers)
```


In this step we will draw a stacked bar lot using ggplot() method having states on X axis and DRIVERS_SPEED and DRIVERS_INVOLVED stacked together on Y axis. To achive this we first used select() method to get required columns. The used gather() method to make data long for DRIVERS_INVOLVED and DRIVERS_SPEED and finally used ggplot() to draw stacked bar plot.

* **select()** function is from **dplyr** package that keeps only the variables we mention.
* **gather()** function is from **tidyr** package that takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed.
* **ggplot()** All **ggplot2** plots begin with a call to ggplot().

```{r}
bad_drivers %>% 
  select(STATE, DRIVERS_INVOLVED, DRIVERS_SPEED) %>% 
  gather(type, value, DRIVERS_INVOLVED:DRIVERS_SPEED) %>% 
  ggplot(., aes(x = STATE,y = value, fill = type)) + 
  geom_bar(position = "stack", stat="identity") + 
  scale_fill_manual(values = c("red", "darkred")) + 
  ylab("Drivers involved in Fatal collision while Speeding") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```

Similarly next stacked plot is having states on X axis and DRIVERS_ALCHO and DRIVERS_INVOLVED stacked together on Y axis. To achive this we first used select() method to get required columns. The used gather() method to make data long for DRIVERS_INVOLVED and DRIVERS_ALCHO and finally used ggplot() to draw stacked bar plot.



```{r}
bad_drivers %>% 
  select(STATE, DRIVERS_INVOLVED, DRIVERS_ALCHO) %>% 
  gather(type, value, DRIVERS_INVOLVED:DRIVERS_ALCHO) %>% 
  ggplot(., aes(x = STATE,y = value, fill = type)) + 
  geom_bar(position = "stack", stat="identity") + 
  scale_fill_manual(values = c("green", "darkgreen")) + 
  ylab("Drivers involved in Fatal collision while Alcho-Impaired") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```


Next stacked plot is having states on X axis and DRIVERS_NOT_DIST and DRIVERS_INVOLVED stacked together on Y axis. To achive this we first used select() method to get required columns. The used gather() method to make data long for DRIVERS_INVOLVED and DRIVERS_NOT_DIST and finally used ggplot() to draw stacked bar plot.



```{r}
bad_drivers %>% 
  select(STATE, DRIVERS_INVOLVED, DRIVERS_NOT_DIST) %>% 
  gather(type, value, DRIVERS_INVOLVED:DRIVERS_NOT_DIST) %>% 
  ggplot(., aes(x = STATE,y = value, fill = type)) + 
  geom_bar(position = "stack", stat="identity") + 
  scale_fill_manual(values = c("lightyellow", "yellow")) + 
  ylab("Drivers involved in Fatal collision not distracted") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```



Next stacked plot is having states on X axis and DRIVERS_NO_ACC and DRIVERS_INVOLVED stacked together on Y axis. To achive this we first used select() method to get required columns. The used gather() method to make data long for DRIVERS_INVOLVED and DRIVERS_NO_ACC and finally used ggplot() to draw stacked bar plot.


```{r}
bad_drivers %>% 
  select(STATE, DRIVERS_INVOLVED, DRIVERS_NO_ACC) %>% 
  gather(type, value, DRIVERS_INVOLVED:DRIVERS_NO_ACC) %>% 
  ggplot(., aes(x = STATE,y = value, fill = type)) + 
  geom_bar(position = "stack", stat="identity") + 
  scale_fill_manual(values = c("blue", "darkblue")) + 
  ylab("Drivers involved in Fatal collision no pre accident") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```


Below plot of for STATE vs INS_PREMIUM that used ggplot() method to draw a bar plot.


```{r}
bad_drivers %>% 
  ggplot(., aes(x = STATE,y = INS_PREM)) + 
  geom_bar(position = "stack", stat="identity") + 
  ylab("Car Insurance Premium") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```



### **Conclusion**
Here we discussed various packages and their functions to explore bad drivers dataset. For complete set details refer  (https://www.tidyverse.org/).



**Resources:**

* [https://www.tidyverse.org/packages/](https://www.tidyverse.org/packages/)
* [https://fivethirtyeight.com/features/which-state-has-the-worst-drivers/](https://fivethirtyeight.com/features/which-state-has-the-worst-drivers/)