---
title: "TidyVerseRecipe"
subtitle: "Nilsa Bermudez"
author: "extended by Bonnie Cooper"
date: "4/19/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: zenburn
---


```{r, message=F}
library(tidyverse);
library(reshape)
library(plyr)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this assignment I will create an example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with my selected dataset.”

In light of the current situation across the world, I chose the Bob Ross elements by episode recordset.  

Load the Bob Ross dataset from fivethirtyeight into the RStudio environment:
```{r RossDataSet}
url<-"https://raw.githubusercontent.com/fivethirtyeight/data/master/bob-ross/elements-by-episode.csv"
BobRoss <- read_csv(url)
df <- data.frame(BobRoss)
```

Use the Reshape library's method 'melt' to transform the data set froma wide format to a long format
```{r UseReshapeToGetAllObjects}
Bulkbobr <- reshape::melt(df, id=c("EPISODE","TITLE"))
Bob <- unique(subset(Bulkbobr, value == 1, select = c("EPISODE","TITLE","variable")))
Bob <-Bob[order(Bob$EPISODE),]
Bob <- rename(Bob, c("variable"="object"))
head(Bob,10)
```

```{r CountFrequenciesOfObjects}
Bob2 <-count(Bob,vars="object", wt_var = NULL)
Bob2 <- Bob2 %>% dplyr::arrange( freq )
```

```{r GetTop10AndDisplay}
  top10 <- tail(Bob2, n = 10)
  ggplot(data=top10,aes(x=object, y=freq)) +  
  ggplot2::aes(x=reorder(object, -freq), y=freq) +
  ggplot2::geom_bar(stat="identity",fill="steelblue") +
  ggplot2::labs(x="Objects",y="Frequency") +
  ggplot2::ggtitle("Most used objects in Bob Ross Paintings") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))
```

Now use tidyr's 'pivot_longer()' and dplyr methods to replicate same reshaping of the the data and reproduce the same figure as above:
```{r useTidyr2Reshape, message=F}
Bob3 <- df %>%
    pivot_longer(cols=3:69,names_to = "variable", values_to = "value") %>%
    filter(value==1) %>%
    select(-value) %>%
    dplyr::rename( object = variable ) %>%
    group_by( object ) %>%
    dplyr::summarise('count'= n()) %>%
    dplyr::arrange( desc(count)) %>%
    dplyr::top_n(10) %>%
    ggplot(aes(x=object, y=count)) +  
    ggplot2::aes(x=reorder(object, -count), y=count) +
    ggplot2::geom_bar(stat="identity",fill="steelblue") +
    ggplot2::labs(x="Objects",y="Frequency") +
    ggplot2::ggtitle("Most used objects in Bob Ross Paintings") +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))
Bob3
```
