---
title: "forcats::fct_lump - Tidyverse - Data 607"
author: "Justin Hsi"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github

---

```{r setup, include=FALSE}
library(tidyverse)
```

# Forcats - a package with useful functions when working with categorical data 

## When working with a categorical variable that potentially has an enormous amount of levels, you can use forcats::fct_lump() to quickly group less common levels and focus on what levels you think are important

Let's load the rarepepe dataset [(article here).](https://fivethirtyeight.com/features/pepe-the-frog-symbolism-cryptoart-blockchain/)

```{r}
order_all = read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/rare-pepes/ordermatches_all.csv')
```

Let's say you wanted to examine a categorical variable with too many levels. The rarepepe dataset has the names of rarepepe artworks which create many levels. If you tried to do a plot of them vs their summed historical forward asset sales, you might get something unusable like this:

```{r}
f = factor(x = order_all$ForwardAsset)
order_all %>%
  group_by(ForwardAsset) %>%
  summarise(summed_for_quant = sum(ForwardQuantity)) %>% 
  ggplot(aes(x = fct_reorder(ForwardAsset, summed_for_quant), y = summed_for_quant)) + 
  geom_col() + coord_flip() + labs(x='rarepepe name', y='total combined forward assets in all transactions')
```

Part of what makes this visualization not helpful is the overwhelming number of levels. We can reduce this with *fct_lump*, as it will group the factors that do not occur frequently enough into a single level called "Other". Here we choose the 15 most frequently occuring levels/artworks :

```{r}
f = factor(x = order_all$ForwardAsset)
order_all$ForwardAsset = fct_lump(f, 15)
order_all %>%
  group_by(ForwardAsset) %>%
  summarise(summed_for_quant = sum(ForwardQuantity)) %>% 
  ggplot(aes(x = fct_reorder(ForwardAsset, summed_for_quant), y = summed_for_quant)) + 
  geom_col() + coord_flip() + labs(x='rarepepe name', y='total combined forward assets in all transactions')
```

## Conclusions
Now we can clearly see the scale of the most transacted rarepepe, PEPEONECOIN, compared to every other rarepepe. This could be a useful first step in analysis, such as which rarepepes you might want to acquire at some point for their value.
Some other things to note include the proportion method, rather than choosing the top n most frequent levels. You may consider using this function with fct_inorder, fct_infreq, or fct_inseq which are related to factor sorting. And finally, there is a similar function fct_lump_min, which has a slightly different functionality when you want to specify which level is the lowest count level before grouping to other.

# Extend

We will now examine some of the other functions of the forcats library that were not included in the original vignette: fct_relevel and fct_infreq.

## fct_infreq

fct_infreq is a tool used to reorder factors by their frequency (or count). It is a more specific use of the fct_reorder tool, and allows you to quickly reorder a factor by its count without creating a count variable. This is applicable to the above example. I also use the helper function fct_rev to reverse the factors and display them in decreasing order rather than increasing order.

```{r}
f = factor(x = order_all$ForwardAsset)
order_all$ForwardAsset = fct_lump(f, 15)
order_all %>%
  ggplot(aes(x = fct_rev(fct_infreq(ForwardAsset)))) + 
  geom_bar() + coord_flip() + labs(x='rarepepe name', y='total combined forward assets in all transactions')
```

This is slightly different to the above example as instead of measuring the sum of ForwardAsset's we are merely measuring the number of times something appeared as a ForwardAsset. However, we are able to quickly generate and display this information with the fct_infreq function.

## fct_relevel

```{r}
#Limit Forward Asset to 3 levels.
order_all$ForwardAsset = fct_lump(f, 5)
levels(order_all$ForwardAsset)
```

We have 3 levels and then our other level. Let us say that we need the factor in a particular order. We can reorder the levels of the factor by using the fct_relevel function.

```{r}
order_all$ForwardAsset <- fct_relevel(order_all$ForwardAsset, 'XCP', 'RAREPEPEPRTY', 'PEPECASH')
levels(order_all$ForwardAsset)
```

The levels of the factor are now in the order we specified. It is also possible to call fct_relevel with a function.

```{r}
order_all$ForwardAsset <- fct_relevel(order_all$ForwardAsset, sort)
levels(order_all$ForwardAsset)
```

by using the function sort as an argument to factor_relevel, we reorder the factors alphabetically. Other potentially useful functions to call would be rev (similar to fct_rev) or sample, which takes a sample of the levels.