---
title: "Tidyverse CREATE"
author: "Bruno de Melo"
date: "3/29/2020"
output: html_document
---


## Video Game sales

I'll demonstrate the power of Tidyverse on a dataset from Kaggle: https://www.kaggle.com/gregorut/videogamesales

Fields in the dataset are:

    Rank - Ranking of overall sales

    Name - The games name

    Platform - Platform of the games release (i.e. PC,PS4, etc.)

    Year - Year of the game's release

    Genre - Genre of the game

    Publisher - Publisher of the game

    NA_Sales - Sales in North America (in millions)

    EU_Sales - Sales in Europe (in millions)

    JP_Sales - Sales in Japan (in millions)

    Other_Sales - Sales in the rest of the world (in millions)

    Global_Sales - Total worldwide sales.


# Data import

Following Wickham & Grolemund data science framework, this is the fist step in any data science project: Import.

```{r}

library(tidyverse)

#importing and converting into a dataframe using piping (%>%), which is widely used in tidyverse.

sales_raw<-read.csv("https://raw.githubusercontent.com/bsvmelo/CUNY/master/vgsales.csv") %>% data.frame()

#quick display
head(sales_raw)

#summary stats
summary(sales_raw)

```

Some statistics:
Number of rows: `r nrow(sales_raw)`    
Number of columns: `r ncol(sales_raw)`   
History size: `r ncol(sales_raw)`   
Number of unique games: `r length(unique(sales_raw$Name))`   
Number of unique platforms: `r length(unique(sales_raw$Platform))`    
Number of unique genres: `r length(unique(sales_raw$Genre))`    
Number of unique publishers: `r length(unique(sales_raw$Publisher))`    
Number of NAs in each column: `r sapply(sales_raw, function(x) sum(is.na(x)))`    
Number of N/A in each column: `r colSums(sales_raw=='N/A')`    

Despite the facts there are no NAs, there are N/A which will have to be dealt with in a later stage.

## Visualization
Following Wickham & Grolemund data science framework, I'll start with some basic visualization using ggplot.


# Top 20 ranked games

```{r}
top_20<-head(sales_raw,20)
# Simple Bar Plot
ggplot(top_20)+ geom_col(mapping=aes(x=reorder(Name,-Global_Sales),y=Global_Sales))+ theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+labs(x="Games", y = "Global Sales in millions")

```



# Platforms
#Number of games sold per platform
```{r}
ggplot(sales_raw) + geom_bar(mapping=aes(x=Platform))+ theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+labs(y = "# units sold in mn")
```

# Genres
# Types of Genres
```{r}
ggplot(sales_raw) + geom_bar(mapping=aes(x=Genre))+ theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+labs(y = "# of games")

```

# Global Sales

# Sales per year per platform
```{r}
ggplot(sales_raw) + geom_col(mapping=aes(Year,Global_Sales,fill=Platform))+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

# Sales per year per platform - Facets

```{r}
ggplot(sales_raw) + geom_col(mapping=aes(Year,Global_Sales,fill=Platform))+facet_wrap(~Platform)+theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position = "none")

```


## Data Transformation

1:
First thing to do is to take care of the N/A values in the dataset.     
To recap, this is the number of N/A in each column: `r colSums(sales_raw=='N/A')`    

There are N/A in the Year and Publisher columns.

I will delete all records in Year and Publisher columns that have an N/A.


```{r}
# replacing N/A by NA
sales_df1<-sales_raw
sales_df1<-sales_df1[sales_df1$Year !="N/A",]
sales_df1<-sales_df1[sales_df1$Publisher !="N/A",]
# confirming that all records containing N/As are deleted
colSums(sales_df1=='N/A')
```


2:Filtering
I will analyse only Sony Platform games, which I assume are: PS, PS2, PS3, PS4, PSP, PSV.

```{r}
# Date frame with only games for Sony platforms
sales_sony<-drop_na(sales_df1)
sales_sony<-filter(na.omit(sales_df1), Platform %in% c('PS', 'PS2', 'PS3', 'PS4', 'PSP', 'PSV'))
head(sales_sony)
```


3:Selecting columns
I will use the select() function to narrow down the analysis to games sold in Japan

```{r}
# Date frame with only sales in Japan
sales_sony_jp<-select(sales_sony, -c(NA_Sales,EU_Sales,Other_Sales))
head(sales_sony_jp)
```

4:Adding new variable with mutate()
I will create a variable corresponding to the Global share of the Japanese sales

```{r}
sales_sony_jp<-mutate(sales_sony_jp,Share=JP_Sales/Global_Sales)
```

5:Grouped Summaries with summarize()
I will summarize the Japanese Share of total sales in aggregate and also by year and publisher.

```{r}
sales_sony_jp_summ_1<-summarize(sales_sony_jp,sum(JP_Sales)/sum(Global_Sales))
sales_sony_jp_summ_2<-sales_sony_jp %>%
  group_by(Year) %>%
  summarize(Share1=sum(JP_Sales)/sum(Global_Sales)) 
```

Historical share of Japanese sales  is: ` r sales_sony_jp_summ_1`


Plotting share of Japanese sales over time:

```{r}
ggplot(sales_sony_jp_summ_2)+ geom_col(mapping=aes(sales_sony_jp_summ_2$Year,sales_sony_jp_summ_2$Share1))+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+labs(y = "JP Share", x="Year")
```


It seems there's an error with 2017 data. I will leave this to my colleagues to try to fix :-)

Aggregation by year and by Platform

```{r}
(sales_sony_jp_summ_3<-sales_sony_jp %>%
  group_by(Year, Platform) %>%
  summarize(Share1=sum(JP_Sales)/sum(Global_Sales))) 
```

Aggregation by year by Genre and by Platform, when Share is greater than 0.

```{r}
(sales_sony_jp_summ_4<-sales_sony_jp %>%
  group_by(Year, Platform, Genre) %>%
  summarize(Share1=sum(JP_Sales)/sum(Global_Sales))) %>%
  filter(Share1>0)

```

Aggregation by Platform, when Share is greater than 0.

```{r}
(sales_sony_jp_summ_5<-sales_sony_jp %>%
  group_by(Platform) %>%
  summarize(Share1=sum(JP_Sales)/sum(Global_Sales))) %>%
  filter(Share1>0)

```


## Conclusion

I have used a lot of different functions to transform, tidy, and present data. There's still a lot analysis that can be done.


