---
title: "Tydiverse_create"
author: "Dennis Pong"
date: "3/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Data

I'll be using the data from [SARS 2003 Outbreak Complete Dataset](https://www.kaggle.com/imdevskp/sars-outbreak-2003-complete-dataset) from Kaggle. The raw github link is [here](https://raw.githubusercontent.com/metis-macys-66898/data_607_sp2020/master/sars_2003_complete_dataset_clean.csv)


## Loading Packages 

```{r}
library(tidyverse)
library(tidyr)
```

## Loading the data 

```{r}
df=read.csv('https://raw.githubusercontent.com/metis-macys-66898/data_607_sp2020/master/sars_2003_complete_dataset_clean.csv', stringsAsFactors	
 = F)

head(df)
str(df)
```

```{r}
as.tibble(df)

```


## Using dplyr: rename to rename variables by name

### Usage
rename(.data, ...)

select(.data, ...)

### Examples

#### rename( ) can be used to rename columns, except all columns not mentioned are kept

```{r rename}
df1 <- df %>% rename("Cumulative_number_of_cases" = "Cumulative.number.of.case.s.")
names(df1)
```

#### Select( ) can be used to rename columns, though all columns not mentioned are dropped
```{r}
df %>% select("Cumulative_number_of_cases" = "Cumulative.number.of.case.s.") %>% head(5)
```

#### Need to make sure df1 is not tibble.

```{r convert_df1_to_tibble, include=FALSE}
df1 <- as.tibble(df1)
```


## Using dplyr: top_frac to select top (or bottom) n rows (by value)

### Usage
top_frac(x, n, wt)

top_n(x, n, wt)  
  
  
### Examples

#### Wanted to get the top 30% of the countries by day in terms of cumulative number of cases
```{r}
# The funciton count will give us a count of records by day, denoted in n. 
# group_by (x) followed by %>% will give you the result grouping by x
df1 %>% group_by (Date) %>% count

```

#### .3 * n( ) will give you 30% of row count. Thus the following will give you the top 30% of the countries per date in terms of cumulative number of cases.
```{r n_30%}
#  Assign the result to df2  
df2 <- df1 %>% group_by(Date ) %>% top_n(.3 * n(), Cumulative_number_of_cases)
df2
```


#### top_frac is basically a shorthand to eliminate the need to put in n() for the total row count in that group. Simply put in the fraction as a the parameter. See below.

```{r top_frac}
# note that top_frac takes in .3 instead of .3 * n(). Assign the result to df2a
df2a <- df1 %>% group_by(Date ) %>% top_frac(.3, Cumulative_number_of_cases)
df2a 

# checking if df2a is the same as df2
all.equal(df2, df2a)
```

  
  
  
### Use dplyr::tally or dplyr::count to do counts like n( )
#### Besides summarize/summarise, what're some of the functions that one can use to tally counts?

### Usage
tally(x, wt = NULL, sort = FALSE, name = "n")

count(x, ..., wt = NULL, sort = FALSE, name = "n", .drop = group_by_drop_default(x))  
  
  
#### tally the record/obs count by country 

```{r tally}
# an automatic variable n is created to hold the results of tally
df2 %>% group_by(Country) %>% tally
```


#### Count the number of record/obs by country 

```{r count}
# It can be further simplified to using just count. count (x) means counting the number of records grouping by x. 
df2 %>% count (Country)

```


