---
title: "forcats::fct_lump - Tidyverse - Data 607"
author: "Justin Hsi"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github

---

```{r setup, include=FALSE}
library(tidyverse)
```

# Forcats - a package with useful functions when working with categorical data 

## When working with a categorical variable that potentially has an enormous amount of levels, you can use forcats::fct_lump() to quickly group less common levels and focus on what levels you think are important

Let's load the rarepepe dataset [(article here).](https://fivethirtyeight.com/features/pepe-the-frog-symbolism-cryptoart-blockchain/)

```{r}
order_all = read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/rare-pepes/ordermatches_all.csv')
```

Let's say you wanted to examine a categorical variable with too many levels. The rarepepe dataset has the names of rarepepe artworks which create many levels. If you tried to do a plot of them vs their summed historical forward asset sales, you might get something unusable like this:

```{r}
f = factor(x = order_all$ForwardAsset)
order_all %>%
  group_by(ForwardAsset) %>%
  summarise(summed_for_quant = sum(ForwardQuantity)) %>% 
  ggplot(aes(x = fct_reorder(ForwardAsset, summed_for_quant), y = summed_for_quant)) + 
  geom_col() + coord_flip() + labs(x='rarepepe name', y='total combined forward assets in all transactions')
```

Part of what makes this visualization not helpful is the overwhelming number of levels. We can reduce this with *fct_lump*, as it will group the factors that do not occur frequently enough into a single level called "Other". Here we choose the 15 most frequently occuring levels/artworks :

```{r}
f = factor(x = order_all$ForwardAsset)
order_all$ForwardAsset = fct_lump(f, 15)
order_all %>%
  group_by(ForwardAsset) %>%
  summarise(summed_for_quant = sum(ForwardQuantity)) %>% 
  ggplot(aes(x = fct_reorder(ForwardAsset, summed_for_quant), y = summed_for_quant)) + 
  geom_col() + coord_flip() + labs(x='rarepepe name', y='total combined forward assets in all transactions')
```

## Conclusions
Now we can clearly see the scale of the most transacted rarepepe, PEPEONECOIN, compared to every other rarepepe. This could be a useful first step in analysis, such as which rarepepes you might want to acquire at some point for their value.
Some other things to note include the proportion method, rather than choosing the top n most frequent levels. You may consider using this function with fct_inorder, fct_infreq, or fct_inseq which are related to factor sorting. And finally, there is a similar function fct_lump_min, which has a slightly different functionality when you want to specify which level is the lowest count level before grouping to other.


