---
title: "forcats makes plots better"
author: "Catlin"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(forcats)  # tidyverse package for factors; needs to be explicitly loaded
```

# load information from fivethirtyeight.com on college majors
Thank you to Ait Elmouden Abdellah for introducing me to this dataset!

```{r}
CollegeMajors <- read.csv("https://raw.githubusercontent.com/aaitelmouden/DATA607S2020/master/Week1/recent-grads.csv")
```

```{r}
str(CollegeMajors)
```

```{r}
str(CollegeMajors)
```

## consider default ordering

```{r}
CollegeMajors %>%
  count(Major_category, wt = Total, sort = TRUE) %>%
  ggplot(aes(Major_category, n)) + geom_col() + coord_flip() +
labs(x = "",
     y = "Total # of Graduates")
```

Consider: 

(1) what is the current sort order?
(2) What is the desired sort order?

```{r}
CollegeMajors %>%
  count(Major_category, wt = Total, sort = TRUE) %>%
  ggplot(aes(x = fct_reorder(Major_category, n), y = n)) + geom_col() + coord_flip() +
labs(x = "",
     y = "Total # of Graduates")
```

Notes: 

* Sometimes, you'll want to wrap the call to fct_reorder() in an fct_rev()
* fct_infreq() can be useful instead of fct_reorder when you're working with a single variable
