---
title: "Data 607 - Bar Chart Visualization Using geom_col()"
author: "Layla Quinones"
date: "3/12/2020"
output: rmdformats::readthedown
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(RCurl)
```

**By:** Layla Quinones

**Date:** 03/12/2020


> Note: tidyverse, Rcurl, kableExtra, colorBrewer, and rmdformats used. 

# I. Abstract

`ggplot2` is a ubiquitous library that allows you to easily visualize data and explain patterns in data to people who are not nessesarliy familiar with the technical aspect of data analysis. Visualizations made with `ggplot2` are easy to understand and contruct thanks to an API that allows visualizations to be "built" via layering of graphics and other visual elements.  In this document we explore the a few ways we can construct bar charts to visualize categorical data, specifically using `geom_col()`. All other functions of `ggplot2` can be found at this [reference link](https://ggplot2.tidyverse.org/reference/).

# II. About the Data

The data that I chose for this presentation provides information about the Incomes of U.S. Workers by Occupation and Gender. This dataset really appealed to me because I have heard some debate about the topic of gender and unequal pay, and I wanted to motivate data analysis and visualization so that it is easy for skeptics to interpret.

The following variables are included:

`Occupation:` Job title as given from BLS. Industry summaries are given in ALL CAPS.

`All_workers:` Number of workers male and female, in thousands.

`All_weekly:` Median weekly income including male and female workers, in USD.

`M_workers:` Number of male workers, in thousands.

`M_weekly:` Median weekly income for male workers, in USD.

`F_workers:` Number of female workers, in thousands.

`F_weekly:` Median weekly income for female workers, in USD.

**DATASET SOURCE**: [Kaggle.com](https://www.kaggle.com/jonavery/incomes-by-career-and-gender#inc_occ_gender.csv)

## Load the Data
```{r}
url <- "https://raw.githubusercontent.com/MsQCompSci/data/master/inc_occ_gender.csv"

rawData <- getURL(url)
rawData <- read.csv(text = rawData, header = TRUE)

kable(rawData)%>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

# III. Visualize the Data Using GGPLOT2 Layers

The basic anotomy of ggplot2 grammer consists of **raw data** that has been fine tuned based on your focus, **layers** which include all the things that are visible about your graph (ie. data points, lines, bars, area shading, title, legend, etc), **scale** which stretches or shrinks the data according to your plot area, **coordinate system** which can be a 2d grid (but others are also possible with ggplot2) and **themes** which includes font, fontsize, color, etc (aesthetic component). The more you want to fine tune your visualization, the more detailed your ggplot2 grammer will be. 

### Context

Before you visualize any data it is a good idea to get to know the dataset. For this data set it would be interesting to compare the mean weekly income for all workers in an occupation with the mean weekly income for females and for males. For this comparison, I will choose a bar chart because it can immediatly give my audience an intuitive idea of what the data is tryin to show.

### Format the data

Now that we have an idea about what we want to visualize and *HOW* we want to visualize it, we can format our data in long format and filter out any data you do not need. We will tidy data into a long format so that ggplot2 knows how to organize that data. To do this we complete the following steps:

1. Select only the columns of data that report weekly earnings using the `select()` function a put them in a new variable

2. Put this new data in long format using `pivot_longer()`

3. Convert all salary values from factors, to characters to numeric (must be done in that order to retain numeric value)

4. Drop all `NA` values using `drop_na()`

```{r, warning = FALSE}
# create a new dataframe called weekly data
weeklyData <- rawData %>% 
  select(Occupation, All_weekly, M_weekly, F_weekly)%>% # 1
  pivot_longer(All_weekly:F_weekly, names_to = "salaryCat", values_to = "salary") #2

#3
weeklyData$salary <- as.numeric(as.character(weeklyData$salary))

#4
weeklyData <- weeklyData %>% 
  drop_na()

#display using kable
kable(weeklyData)%>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

## Create a simple bar chart

First we will focus on a simple bar chart that shows the weekly salaries for male and females from all professions. In `ggplot2` we can use two functions to create bar charts: `geom_bar()` and `geom_col()`. Here we will be using `geom_col()` because it will allow us to map bar size to one of our variables (salary). `geom_bar()` is best used when you are creating bar charts that visualize frequency. 

To create a simple bar chart using `geom_col()` we need to do the following:

1. Filter the data that corresponds to all professions using the `filter()` function to select only the data that corresponds to rows with `ALL OCCUPATIONS` data.

2. Begin our chart by using `ggplot()` function and designating the dataframe we want to visualize

3. Lay on the first layer by first adding the `+` symbol and calling on the`geom_col()` function which create a bar chart. 

4. In `geom_col()` we must define which columns from our dataframe will be along the x axis and y axis using the `aes()` function.

```{r}
#1
allOcc <- weeklyData %>%
  filter(Occupation == "ALL OCCUPATIONS")

ggplot(allOcc) + #2
  geom_col(aes(x = salaryCat, y = salary)) #3 & 4
```

Now that we have our first simple bar chart we can layer on various components to make this graph more visually appealing. The first thing that comes to mind is to order the bars according to salary. We do this by adding the `reorder()` command to the x value in `aes()`.

```{r}
 ggplot(allOcc) +
  geom_col(aes(x = reorder(salaryCat, salary), y = salary))
```

Then to add color to each category we can use the `fill` attribute **AFTER** `aes()`. We can choose colors by defining a character vector with specific color values for each category `fill = c("magenta", "lightgreen", "lightblue")`. 

```{r}
 ggplot(allOcc) +
  geom_col(aes(x = reorder(salaryCat, salary), y = salary), fill = c("magenta", "lightgreen", "lightblue"))
```

To change the labels for each category so that they are more readable to viewers we can add the next layer using `scale_x_discrete(labels = "Female", "ALL", "Male"))`. NOTE: between each layer always include the `+` symbol.

```{r}
ggplot(allOcc) +
  geom_col(aes(x = reorder(salaryCat, salary), y = salary), fill = c("magenta", "lightgreen", "lightblue")) +
  scale_x_discrete(labels=c("Female", "ALL", "Male")) 
```

To change the x and y titles, as well as the title of the chart we can add another layer using `labs()` and center it in `theme()`. To change the x and y titles, as well as the title of the chart we can add another layer using `labs()` and center it in `theme()` by setting the `plot.title` attribute and `element_text()` function to adjust the position using `hjust` to the middle wich corresponds to 0.5 (for half of the plot width). Also in theme we can change the background by setting the `panel.background` attribute and setting the rectangular panel using `element_rect` to a color using `fill`. I chose the color black here because I like the contrast on bright colors on a dark background

```{r}
ggplot(allOcc) +
  geom_col(aes(x = reorder(salaryCat, salary), y = salary), fill = c("magenta", "lightgreen", "lightblue")) +
  scale_x_discrete(labels=c("Female", "ALL", "Male")) +
  labs(y = "Gender", x ="Salary", title = "Mean Weekly Salary for US Professionals") +
  theme(plot.title = element_text(hjust = 0.5), panel.background = element_rect(fill = "black"))
```

For the viewers' convenience I would like to include the value of the salary on top of each bar. To do this we layer on the `geom_text` function which will add text to the plot. To place the text on top of each bar the x value needs to correspond to the category of `salary`, the y value should be a little below the bar therefore I used `salary - 20`, the actual text should correspond to the value of `salary` which we define using `label`

```{r}
ggplot(allOcc) +
  geom_col(aes(x = reorder(salaryCat, salary), y = salary), fill = c("magenta", "lightgreen", "lightblue")) +
  scale_x_discrete(labels=c("Female", "ALL", "Male")) +
  labs(y = "Gender", x ="Salary", title = "Mean Weekly Salary for US Professionals") +
   theme(plot.title = element_text(hjust = 0.5), panel.background = element_rect(fill = "black")) + 
  geom_text(aes(x = salaryCat, y = salary - 20, label = salary))
```

As we can see here now it is totally obveous that males have a higher mean salary than females in the United States. In addition, the mean salary for males also exceed the mean salary for all genders. 

## Create a grouped bar chart

Now lets say we want to go more into detail about these salaries. We can easily visualize the salaries for each category by created a The weekly salary data is grouped in categories therefore a bar chart will be appropriate to visualize and compare the salaries in the different groups. To do this we complete the following steps:

**1. Filter Data: **

- Filter weekly salary for males and females using `filter()` from the `weeklyData` dataframe

- Conver from long format to wide using `spread()`

- Drop any rows that contain `NA` values (those are not compareable) using `drop_na()`

- Convert back to long format using `pivot_longer()`

**2. Graph Data:**

- Begin chart by using `ggplot()`

- First layer using `geom_col()` for bar chart

- In `geom_col()` define columns for x axis, y axis and color parameter using the `aes()`

**Note:** The big change here is that we are defining `fill` inside of `aes()` and setting it equal to the `salaryCat` column which identifies the group each salary belongs to (male or female)

- Set the colors for male and female bard manually using `scale_fill_manual()`

- Flip the graph so that the bars grow horizontally using `coord_flip()`

- Change the position of the legend using `theme(legend.position = "top")`

```{r}
#1
femaleMale <- weeklyData %>% 
  filter(salaryCat == "F_weekly" | salaryCat == "M_weekly") %>% 
  spread(salaryCat, salary) %>% 
  drop_na() %>% 
  pivot_longer(F_weekly:M_weekly, names_to = "salaryCat", values_to = "salary") 

ggplot(femaleMale) +
  geom_col(aes(x = reorder(Occupation,salary), y = salary, fill = salaryCat),
           position = "dodge") +
  scale_fill_manual(values=c("magenta", "lightblue"))+
  coord_flip() +
  theme(legend.position = "top")
```

As we can see from the plot above, the data seems very bunched up and messy. This is not visually appealing or easily interpretable. What we can do instead is filter the occupations by salary and choose only the top 10 highest salary occupations to analyze. 

**1. Filter the data**

- Convert to wide format using `spread()`

- Filter rows where the column for male weekly salaries exceed \$1500

- Convert back to long format using `pivot_longer()`

- Select the top 10 occupations (first 20 rows)

**2. Graph Data:**

- Begin chart by using `ggplot()`

- First layer using `geom_col()` for bar chart. 

- In `geom_col()` define columns for x axis, y axis and color parameter using the `aes()`

- Set the colors for male and female bard manually using `scale_fill_manual()`

- Flip the graph so that the bars grow horizontally using `coord_flip()`

- Label x and y axis and title using `labs()`

- Change the position of the title and the legend using `theme()`

```{r}
#1
femaleMale <- femaleMale %>% 
  spread(salaryCat, salary) %>% 
  filter(M_weekly >1500) %>%
  pivot_longer(F_weekly:M_weekly, names_to = "salaryCat", values_to = "salary")

femaleMaleHigh <- femaleMale[1:20,]

#2
ggplot(femaleMaleHigh) +
  geom_col(aes(x = reorder(Occupation,salary), y = salary, fill = salaryCat), position = "dodge") +
  scale_fill_manual(values=c("magenta", "lightblue"))+
  coord_flip() +
  labs(y = "Salary (USD)", x ="Occupation", title = "Top 10 Mean Weekly Salary for US Professionals") +
   theme(plot.title = element_text(hjust = 0.5), legend.position = "top") 
```

We can clearly see the discrepencies between male and female salaries for the top 10 highest paying jobs. In addition to the salaries it would also be interesting to use `ggplot2` to visualize the number of males and females in these jobs.

## A Different Take On the Data (Grouped Bar Charts)
```{r, warning = FALSE}
# create a new dataframe with number of workers and convert to long format
workerData <- rawData %>% 
  select(Occupation, M_workers, F_workers)%>%
  pivot_longer(M_workers:F_workers, names_to = "Gender", values_to = "Workers") %>%
  drop_na() 

# Select only the rows that match with the top 10 paying jobs
workerData <- workerData %>% filter(Occupation == femaleMaleHigh$Occupation[1] | Occupation == femaleMaleHigh$Occupation[3] | Occupation == femaleMaleHigh$Occupation[5] | Occupation == femaleMaleHigh$Occupation[7] |Occupation == femaleMaleHigh$Occupation[9]|Occupation == femaleMaleHigh$Occupation[11]|Occupation == femaleMaleHigh$Occupation[13]|Occupation == femaleMaleHigh$Occupation[15]|Occupation == femaleMaleHigh$Occupation[17]|Occupation == femaleMaleHigh$Occupation[19])

#display using kable
kable(workerData)%>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

# Using the same method of plotting as above section (grouped bar charts)
ggplot(workerData) +
  geom_col(aes(x = reorder(Occupation,Workers), y = Workers, fill = Gender), position = "dodge") +
  scale_fill_manual(values=c("magenta", "lightblue"))+
  coord_flip() +
  labs(y = "Number of Workers", x ="Occupation", title = "Number of Workers for the Top 10 Paying Jobs") +
   theme(plot.title = element_text(hjust = 0.5), legend.position = "top") 
```

Not surprsingly there is a **HUGE** difference in the number males and females in fields designated as `COMPUTATIONAL`. Through this visualization, it is easy to provide evidence for the gender divide, unequal pay and gender bias in high paying jobs. 