---
title: "Working with Lubridate"
author: "Ken Popkin"
date: "3/12/2020"
output: html_document
---

#### Dates are a frequent feature in data analysis and data science projects.  In this vignette we'll look at the Lubridate package and perform a few date transformations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
```

#### Load information from Kaggle's Hourly Energy Consumption dataset
[link to data description](https://www.kaggle.com/robikscube/hourly-energy-consumption/data#)

```{r}
energy <- read.csv('C:/Users/user/Documents/00_Applications_DataScience/CUNY/DATA607/DATA607-Class-Repository/zTidyVerse_Assignment/DEOK_hourly.csv')

energy2 <- head(energy, 1)
energy2
```


#### Now we'll explore some of the functionality that makes Lubridate so helpful and expand the date features in the energy dataframe.

##### Parsing enables various parts of a date to be obtained
```{r}
#get the year, month and day
energy2$year <- year(energy2$Datetime)
energy2$month <- month(energy2$Datetime)
energy2$day <- day(energy2$Datetime)
```

```{r}
#get the week of the year and the iso week of the year
energy2$week <- week(energy2$Datetime)
energy2$isoweek <- isoweek(energy2$Datetime)

select(energy2, Datetime, year, month, day, week, isoweek)
```


##### Timezones are another convenient and easy to obtain feature
```{r}
#Uncomment to see over 600 time zones that R can provide:   OlsonNames()

energy2$Datetime <- as_datetime(energy2$Datetime)

energy2$atlantic_tz <- with_tz(energy2$Datetime,"US/Atlantic")
energy2$pacific_tz <- with_tz(energy2$Datetime,"US/Pacific")

select(energy2, Datetime, atlantic_tz, pacific_tz)
```


##### Lubridate simplifies math operations on dates
```{r}
#add and subtract some days to Datetime and create a new feature
energy2$two_weeks_out <- energy2$Datetime + days(14)

energy2$two_weeks_back <- energy2$Datetime + days(-14)

select(energy2, Datetime, two_weeks_back, two_weeks_out)
```


#### Conclusion: 
This has been a small subset of the many capabilities of the Lubridate package.  There are a number of other capabilities that could be added to this vignette.
```{r}
energy2
```
