---
title: "TidyVerseRecipe"
author: "Nilsa Bermudez"
date: "3/27/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: zenburn
---


```{r}

library(tidyverse);
library(reshape)
library(plyr)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this assignment I will create an example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with my selected dataset.”

In light of the current situation across the world, I chose the Bob Ross elements by episode recordset.  


```{r RossDataSet}
url<-"https://raw.githubusercontent.com/fivethirtyeight/data/master/bob-ross/elements-by-episode.csv"
BobRoss <- read_csv(url)
df <- data.frame(BobRoss)

```

```{r UseReshapeToGetAllObjects}

Bulkbobr <- reshape::melt(df, id=c("EPISODE","TITLE"))
Bob <- unique(subset(Bulkbobr, value == 1, select = c("EPISODE","TITLE","variable")))

Bob <-Bob[order(Bob$EPISODE),]
Bob <- rename(Bob, c("variable"="object"))
head(Bob,10)
```


```{r CountFrequenciesOfObjects}
Bob2 <-count(Bob,vars="object", wt_var = NULL)

```

```{r GetTop10AndDisplay}
  top10 <- head(dplyr::arrange(Bob2), n = 10)
  ggplot(data=top10,aes(x=object, y=freq)) +  
  ggplot2::aes(x=reorder(object, -freq), y=freq) +
  ggplot2::geom_bar(stat="identity",fill="steelblue") +
  ggplot2::labs(x="Objects",y="Frequency") +
  ggplot2::ggtitle("Most used objects in Bob Ross Paintings") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))

```





