---
title: "Tidyverse Create"
author: "Ait elmouden"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Objectives

The goal of this assignment is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com, we'll create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

### Source of Data 

We will use the fivethirtyeight Package, so we can load any data we want from the R package instead of through the importing of comma separated values (CSV) files, Excel spreadsheets, andother file formats.

First we'll load the fivethirtyeight package, as well as tidyr, tibble and dplyr and then call murder_2015_final

Let's explore the data

```{r}
library(fivethirtyeight)
library(tidyr)
library(tibble)
library(dplyr)

murder_data <- murder_2015_final
head(murder_data)
```

### Look at the dataset’s column names

Use names(Data) to list out the dataset’s column names.

```{r}
names(murder_data)
```


### Gather variables into a single column

Lets gather up the two year variables, murders_2014 and murders_2015 into a single year column we’ll name murder_year. We’ll store the number of murders in a column titled murders and call this new object murders_gathered.

```{r}
murders_gathered <- murder_data %>% 
    gather(
        murder_year,
        murders,
        murders_2014:murders_2015,
        na.rm = TRUE)
murders_gathered
```

### Arrange data alphabetically by state and city

```{r}
murders_arranged <- murders_gathered %>% 
    arrange(
        state, 
        city)
murders_arranged
```

So now we have the two years in a single column (murder_year), but the repeating murder_ text is redundant. We want to get the year in a column by itself.

### Separate “murder_year” column into “text” and “year”

```{r}
murders_separate <- murders_arranged %>%
    separate(
        murder_year,
            into = c("text", 
                     "year")
        )
murders_separate
```

Now we can use spread() to put the year back into two different columns, 2014 & 2015. But I will combine this with arrange() so the output is easier to read.

```{r}
murders_spread <- murders_separate %>% 
    spread(
        year,
        murders
        ) %>% 
        arrange(
            state,
            city)
murders_spread
```

### Combine city and state into a single column city_state using unite()

```{r}
murders_final <- murders_spread %>%
unite(
    city_state, 
    city, 
    state) %>% 
        arrange(
            city_state
            ) %>% 
            select(
                -(text)
                )
murders_final
```

### Output the new table as a csv

```{r}
write.csv(murders_final, file = "murders_final.csv",row.names=FALSE, na="")
```

