---
title: "Data Acquisition and Management Week 9 Assignment - TidyVerse"
author: "Chun San Yip"
date: "2020/03/29"
output: html_document
---

Overview:
The task here is to create an example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

The core tidyverse packages includes the following packages:
1. ggplot2
2. dplyr
3. tidyr
4. readr
5. purr
6. tibble
7. stringr
8. forcats

I will focus on the packages dplyr in this exercise, there are close to 70 functions in the packages dplyr. I will focus on the following functions for Basic single-table:
1. select(), rename() - Select/rename variables by name
2. arrange() - Arrange rows by variables
3. filter() - Return rows with matching conditions
4. pull() - Pull out a single variable
5. mutate() transmute() - Create or transform variables
6. summarise() summarize() - Reduce multiple values down to a single value
7. group_by() ungroup() - Group by one or more variables

Load tidyverse package
```{r load-packages, message = FALSE}
library(tidyverse)
```

We need to read a CSV file from FiveThirtyEight.com. The Raptor data is a ratings which uses play-by-play and player tracking data to calculate each player's individual plus-minus measurements and wins above replacement, which accounts for playing time. Function read_csv can be used to read the CSV file and convert in a dataframe
```{r}
theUrl1 <- "https://projects.fivethirtyeight.com/nba-model/2020/latest_RAPTOR_by_team.csv"

raptor <- read_csv(file=theUrl1,  na = c("", "NA"))


#Table raptor is a dataframe
raptor
```


#1. As there are too many columns, I only want to look at basic data and column with raptor data.It will drop all war, predator and pace columns.
```{r}
raptor1 <- raptor %>%
  select(player_name:raptor_total)

raptor1

```

#2. The next is arrange, I would like to arrange the dataframe by player_name
```{r}
#The next function is arrange
raptor1 <- raptor1 %>%
  arrange(player_name)

raptor1
```

#3. The next is filter, I would like to look at players who play more than 2000 minutes in the season
```{r}
#The next function is arrange
raptor1 <- raptor1 %>%
  filter(mp>=2000)

raptor1
#There are only 27 players
```

#4. The next function is pull, I would like to look at all the player names for these "high usage" players
```{r}
#The next function is arrange
raptor2 <- raptor1 %>%
  pull(1)

raptor2
```

#5. The next function is mutate and/or transmute. As transmute will drop an existing variables, I will use mutate instead. I want to create a new column for a ratio between raptor_box_total and raptor_onoff_total
```{r}
raptor1 <- raptor1 %>%
  mutate(ratio=raptor_box_total/raptor_onoff_total)

raptor1
```

#6 and 7. I would like to use group by and then summarise to look at the average raptor_box_total by teams.
```{r}
raptor1 <- raptor1 %>%
  group_by(team) %>%
  summarise(mean(raptor_box_total))

raptor1
```


***

### Extend Assignment - Leo Yi

First, I'd like to admit that I'm not actually a basketball fan so these calculations aren't very clear to me. From what I can tell, the raptor dataset is an evalulation of individual players based on different metrics. 

The code above, by Chun San Yip, seems to ulitmately summarize the average box total by team, for players who have at least 2,000 minutes of play in the 2020 season.

I'll be using the tidyverse package to create a few visualizations using ggplot2 to try and understand what's being calculated above.

#### 2,000 Minutes of Play Time

The first thing I want to take a look at is why 2000 minutes of play time was used above. Lets use ggplot to see the distribution of play time of each of the players in the raptor dataset and see how many players go into each team's overall average box total.

```{r}
# plot a histogram of each players' active play time
ggplot(raptor, aes(x = mp)) +
  geom_histogram() +
  theme_bw() +
  labs(title = 'Histogram of Minutes Played',
       x = element_blank(),
       y = element_blank())
  
```

It looks like we're only looking at a very small fraction of the players in this dataset. Let's determine how many, exactly. Let's do this by creating a flag for the most active players and showing the results, using dplyr:

```{r}
# add a flag for players with 2000 or more minutes of playtime
raptor_aflag <- raptor %>%
  mutate(active_flag = ifelse(mp >= 2000, 1, 0))

# calculate the percent of the most active players
raptor_aflag %>%
summarize(active_players = sum(active_flag),
          total_players = n(),
          pct_active_players = active_players / total_players)
```

It seems like we're barely looking at 5% of the players in this dataset. This makes for a curious team level average, as it seems possible that a few, perhaps even only one player, is making up the team average calculated above. Lets use dplyr again to take a look at how many active players there are per team:

```{r}
# show the number of active players per team
team_pcts <- raptor_aflag %>%  
  group_by(team) %>%
  summarize(active_players = sum(active_flag),
            total_players = n(),
            pct_active_players = active_players / total_players) %>%
  arrange(desc(pct_active_players))

head(team_pcts)

# show the number of active players per team
ggplot(team_pcts, aes(x = active_players)) +
  geom_bar() +
  scale_y_continuous(limits = c(0,12),
                     breaks = seq(0,12,1)) +
  labs(title = 'Distribution of Active Players by Teams',
       x = "Number of Active Players",
       y = "Number of Teams") +
  theme_bw()

# table version of visualization
team_pcts %>%
  group_by(active_players) %>%
  summarize(number_of_teams = n(),
            pct_of_grp = n() / nrow(team_pcts),
            min_pct_active_players = min(pct_active_players),
            max_pct_active_players = max(pct_active_players))
```

#### Results

It looks like 11 teams or over a third of all the teams listed here won't be included in the average team box totals above. Also, 12 teams, or 40% of the teams in the dataset, are averaged using only one player on their team. Even with 3 players with over 2,000 minutes of play time in one team, we're only looking at 15% of that team's players.

I'm still unclear as to what this metric is and why it's important to basketball-- but it suggests interest in the most active players rather than the combination of the team. Perhaps it would be better to look at box scores by player rather than by team.



