---
title: "Data Acquisition and Management Week 9 Assignment - TidyVerse"
author: "Chun San Yip"
date: "2020/03/29"
output: html_document
---

Overview:
The task here is to create an example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

The core tidyverse packages includes the following packages:
1. ggplot2
2. dplyr
3. tidyr
4. readr
5. purr
6. tibble
7. stringr
8. forcats

I will focus on the packages dplyr in this exercise, there are close to 70 functions in the packages dplyr. I will focus on the following functions for Basic single-table:
1. select(), rename() - Select/rename variables by name
2. arrange() - Arrange rows by variables
3. filter() - Return rows with matching conditions
4. pull() - Pull out a single variable
5. mutate() transmute() - Create or transform variables
6. summarise() summarize() - Reduce multiple values down to a single value
7. group_by() ungroup() - Group by one or more variables

Load tidyverse package
```{r load-packages, message = FALSE}
library(tidyverse)
```

We need to read a CSV file from FiveThirtyEight.com. The Raptor data is a ratings which uses play-by-play and player tracking data to calculate each player's individual plus-minus measurements and wins above replacement, which accounts for playing time. Function read_csv can be used to read the CSV file and convert in a dataframe
```{r}
theUrl1 <- "https://projects.fivethirtyeight.com/nba-model/2020/latest_RAPTOR_by_team.csv"

raptor <- read_csv(file=theUrl1,  na = c("", "NA"))


#Table raptor is a dataframe
raptor
```


#1. As there are too many columns, I only want to look at basic data and column with raptor data.It will drop all war, predator and pace columns.
```{r}
raptor1 <- raptor %>%
  select(player_name:raptor_total)

raptor1

```

#2. The next is arrange, I would like to arrange the dataframe by player_name
```{r}
#The next function is arrange
raptor1 <- raptor1 %>%
  arrange(player_name)

raptor1
```

#3. The next is filter, I would like to look at players who play more than 2000 minutes in the season
```{r}
#The next function is arrange
raptor1 <- raptor1 %>%
  filter(mp>=2000)

raptor1
#There are only 27 players
```

#4. The next function is pull, I would like to look at all the player names for these "high usage" players
```{r}
#The next function is arrange
raptor2 <- raptor1 %>%
  pull(1)

raptor2
```

#5. The next function is mutate and/or transmute. As transmute will drop an existing variables, I will use mutate instead. I want to create a new column for a ratio between raptor_box_total and raptor_onoff_total
```{r}
raptor1 <- raptor1 %>%
  mutate(ratio=raptor_box_total/raptor_onoff_total)

raptor1
```

#6 and 7. I would like to use group by and then summarise to look at the average raptor_box_total by teams.
```{r}
raptor1 <- raptor1 %>%
  group_by(team) %>%
  summarise(mean(raptor_box_total))

raptor1
```
