---
title: "***stringr:*** a package used to manipulate strings"
author: "Devin Teran"
date: "2020-03-28"
output: rmdformats::readthedown

---
### **Getting started**

First we need to load these packages:

* **tidyverse** 
* **stringr**  
* **dplyr** - *used for subsetting data in our analysis*  
* **rmdformats** - *used to for styling html document*   

```{r necessary-libraries,include = FALSE}
packages <- c('tidyverse','stringr','dpylr','rmdformats')

library(tidyverse)
library(stringr)
library(rmdformats)
library(dplyr)
```
  
We're going to load a dataset from [fivethirtyeight.com](https://fivethirtyeight.com/features/a-handful-of-cities-are-driving-2016s-rise-in-murders/) to help us show examples of stringr at work. Our data shows murders in cities in America from 2014 to 2015. 

We'll take the first 10 rows of the data for simplicity's sake.
```{r setup}
url <- 'https://raw.githubusercontent.com/fivethirtyeight/data/master/murder_2016/murder_2015_final.csv'
murder_raw <- read_csv(url)

murder <- head(murder_raw,10)
```

## **Ordering Strings**
#### **str_order(character vector,decreasing = X)**  

**Purpose:**   
Order a character vector alphabetically.  

**Input: **      
character vector - what you want to order  
X - indicate whether to order characters decreasing (FALSE - alphabetically) or increasing (TRUE - order from Z to A)  

**Output:**  
An ordered character vector 

**Example:**  
We'll order the column 'city' from our dataframe 'murder'
```{r sort-descreasing-alphabetically}
str_sort(murder$city,decreasing=FALSE)
```
If you want to reverse the order to Z-A you can set **decreasing = FALSE**
```{r sort-reverse}
str_sort(murder$city,decreasing=TRUE)
```
   
## **Combining Strings**
#### **str_c(String1,String2,...Stringn)**  

**Purpose:**    
The function takes in a strings or vectors of strings and concatentates them together
  
**Input:**     
String or vector of strings separated by comma  
  
**Output:**  
Single string of vector of combined strings  
  
**Example:**  
You can combine as many strings as you want together at once  
```{r combine-many-string}
str_c('a','b','c','d','e','f','g','h')
```  
Let's let's see how we can combine two vectors of strings together from our dataframe: the city and the state  
```{r combine-city-state}
str_c(murder$city,murder$state)
```
You can add a separator between the strings you're combining using the **sep = ''** argument. Let's separate the city and state by a comma.  
  
Add this new data as a column, named **City_State**, in our dataframe murder.
```{r combine-string-separated}
murder$City_State <- str_c(murder$city,murder$state,sep=",")
murder$City_State
```
 
## **Replacing Strings** 
#### **str_replace_all(string, pattern, string)**
**Purpose:**    
This function will replace all instances of a pattern with the given replacement  

**Input:**      
String or vector of strings    
Pattern - you can use [regular expressions](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) here

**Output:**  
Single string of vector of combined strings  

**Example:**  
Supposed we wanted to replace all appearances of **.** in the column 'City_State'.  We can easily do this with **str_replace_all()** 
```{r string-replace-all}
murder$City_State <- str_replace_all(murder$City_State,'[\\.]','*')
murder$City_State
```
   
## **Get the Length of a String**
#### **str_length(string)**
**Purpose:**    
Find out the length of a string or a vector of strings   
  
**Input:**       
String or vector of strings     

**Output:**     
Integer    
  
**Example:**  
Let's find how out how long each city name   
```{r string-length}
str_length(murder$city)
```
  
*Let's only view the rows in the dataframe where the city has more than 9 letters in the name.  To do this we'll also use the filter function from the package dplyr.*  
```{r view-more-9-letter-city}
filter(murder,str_length(murder$city) > 9)
```

### **Conclusion**
These examples are just the beginning of what you can do with stringr.  If you need to manipulate, combine or work with strings in general, stringr is a great package to do so.  Here's a great stringr cheatsheet released by RStudio (https://rstudio.com/resources/cheatsheets/).



**Resources:**

* [https://cran.r-project.org/web/packages/rmdformats/vignettes/introduction.html](https://cran.r-project.org/web/packages/rmdformats/vignettes/introduction.html)
* [https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html)
