---
title: 'DATA 607: Tidyverse Create Assignment'
author: "James Williams"
date: "3/29/2020"
output:
  html_document:
    df_print: paged
---
# Task
Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

# Import Data
Data comes from https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset.
```{r, echo=TRUE, warning = FALSE, message=FALSE}
library(tidyverse)
```
```{r, echo=TRUE, warning = FALSE, message = FALSE}
URL <- 'https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv'
death_rates <- read_csv(URL)
head(death_rates)
```
# Condition Data
```{r, echo=TRUE, warning = FALSE}
death_sums <-death_rates %>%
  replace(is.na(.), 0) %>%
  mutate(sum = rowSums(.[5:71]))
head(death_sums)
```
```{r, echo=TRUE, warning = FALSE}
data<- death_sums[, c("Country/Region", "sum")]
data<- rename(data, Country = "Country/Region", Death_Count = "sum")
head(data)
```
```{r, warning = FALSE}
final_data <- data %>%
  group_by(Country) %>%
  summarize(sum(Death_Count))
head(final_data)
```
```{r, warning = FALSE}
final_data <- rename(final_data, Death_Count = `sum(Death_Count)`)
head(final_data)
```
# Visualize Data
```{r, warning = FALSE}
graph <- ggplot(final_data, aes(x = Country, y = Death_Count)) + 
  geom_point() + 
  theme(axis.text.x = element_text(size = 3, angle = 90))  + 
  geom_text(aes(label=ifelse(Death_Count>5000, as.character(Country), '')), hjust=1, vjust=1) + 
  labs(x = "Country", y = "Death Count", title = "COVID 19 Pandemic: Total Death Counts Worldwide Since 22 January 2020") +
  scale_y_discrete()
graph
```

# Conclusion
Tidyverse features helpful tools to transform messy data sets into more accessible formats.

# Resources
https://www.tidyverse.org/packages/